Article Title

### 2014-12-07 manuscript.tex

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations -- their `functional connectivity' -- is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (`interactions') combined with a low-rank component. The inferred interactions were both positive (`excitatory') and negative (`inhibitory') and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.

@@

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf sample}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \emph{i.e.}\;toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

@@

We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for representing the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. In our data, the sample correlation coefficients were largely positive and low.  We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.  This estimator revealed a sparse network of partial correlations (`{interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons. We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, we found that these interactions exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations. In contrast, the inferred negative interactions were less selective.

@@

Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an unbiased estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
\begin{equation}\label{eq:sample}
    C_{\sf sample} = \frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
    \end{equation}
where $n$ is the number of time bins in $T$. 
The sample covariance matrix yields an unbiased estimate of the true covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean activity, $\bar x$, is estimated independently, in a separate data sample. 
In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the means are estimated independently.

@@

The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}. Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}. For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}. More generally, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}. As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.

Pairwise partial correlations are closely related to the elements of the \emph{precision matrix}, \emph{i.e.}\;the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}. Zero elements in the precision matrix signify zero partial correlation between the two variables. Given the covariance estimate $C$, the matrix of partial correlations $P$ is computed by normalizing the rows and columns of the \emph{precision matrix} $C^{-1}$ to produce negative unit entries on the diagonal:
\begin{equation}\label{eq:partial}
    P = -\left(\mbox{diag}(C^{-1})\right)^{-\frac 1 2} C^{-1} \left(\mbox{diag}(C^{-1})\right)^{-\frac 1 2}
\end{equation}

Increasing the number of recorded neurons results in a higher \emph{condition number} of the covariance matrix \cite{Ledoit:2004} making the partial correlation estimates \emph{ill-conditioned}: small errors in the covariance estimates translate into much greater errors in the estimates of partial correlations. With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization} \cite{Ledoit:2004,Schafer:2005}.

@@

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf sample}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf sample} + \lambda D
\end{equation}
The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (Fig.~\ref{fig:1} Row 1, A).  If sample correlations are largely spurious, $C_{\sf diag}$ is expected to be more efficient than other estimators.

@@

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$:
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \emph{i.e.}\;one in which a large fraction of off-diagonal elements are set to zero.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$. The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1}} Row 1, C).

@@

\subsection*{Simulation}
We next demonstrate how the most efficient among different regularized estimators can reveal the structure of correlations.
We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (Fig.~\ref{fig:1} Row 2, A--D and Methods).  We used these covariance matrices as the ground truth in multivariate Gaussian distributions with zero means and drew samples of various sizes. 
The sample correlation matrices from finite samples (\emph{e.g.}\ $n=500$ in Fig.~\ref{fig:1} Row 3) were contaminated with sampling noise and their underlying structures were difficult to discern.

The evaluation of any covariance matrix estimator, $C$, is performed with respect to a \emph{loss function} $\ell(C,\Sigma)$ to quantify its discrepancy from the truth, $\Sigma$.  The loss function is chosen to attain its minimum when $C=\Sigma$.
Here, in the role of the loss function we adopted the Kullback-Leibler divergence between multivariate normal distributions with equal means, scaled by $\frac 1 p$ to make its values comparable across different population sizes: 
\begin{equation}\label{eq:loss}
    \ell(C,\Sigma) = 
    \frac 1 p D_{KL}\left(\mathcal N_\Sigma\,\Vert\,\mathcal N_C\right) = 
    \frac 1 {2p} \left[\Tr(C^{-1}\Sigma) + \ln\det C - \ln\det\Sigma - p\right] 
\end{equation}
Therefore, $\ell(C,\Sigma)$ is expressed in nats/neuron per time bin.

We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).

@@

In more realistic conditions, when the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
Then the \emph{validation loss} $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 1 p$ and omitting the constant term: 
\begin{equation}\label{eq:vloss}
    \loss{C,C^\prime_{\sf sample}} = \frac 1 {2p} \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
\end{equation}

Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, the validation loss is an unbiased estimate of the true loss, up to a constant:
\begin{equation}
    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
\end{equation}
Therefore, the validation procedure allows comparing the relative values of loss of different covariance estimators.

Indeed, the validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators without access to the ground truth (Fig.~\ref{fig:1} Row 6).

In the example above, the data were sampled from  multivariate normal random variables. In such models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data. 
To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.

Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
Conveniently, the Ising model has equivalent mathematical form to the Gaussian distribution,
\begin{equation}\label{eq:ising}
    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right)
\end{equation}
but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives, although they do not have a simple relationship to the means and the covariance matrix. 
Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.

@@

The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:6} C). This suggests that correlations between neurons build up from multiple chains of smaller interactions. Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting greater uniformity of average partial correlations than the smaller populations ($p=0.002$ Brown-Forsythe test).

@@

\subsection*{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data can produce very large errors in the inferred network of dependencies. The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}. However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be practical or sufficient, and the problem must be addressed by using regularized estimators. Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.

>>>>

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations -- their `functional connectivity' -- is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations arising from linear interactions between pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (`interactions') combined with a low-rank component. The inferred interactions were both positive (`excitatory') and negative (`inhibitory') and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.

@@

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf sample}$] -- the sample covariance matrix, the unbiased estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of covariances toward zero, \emph{i.e.}\;toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- a low-rank approximation of the sample covariance matrix, representing inputs from unobserved shared factors (latent units).
\item[$C_{\sf sparse}$] -- sparse partial correlations, \emph{i.e.}\;a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

@@

We then performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for representing the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. In our data, the sample correlation coefficients were largely positive and low.  We found that the best estimate of the correlation matrix was $C_{\sf sparse+latent}$.  This estimator revealed a sparse network of partial correlations (`interactions'), between the observed neurons; it also inferred latent units exerting linear effects on the observed neurons. We analyzed these networks of partial correlations and found the following: Whereas significant noise correlations were predominantly positive, the inferred interactions had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, the average partial correlations and the inferred positive interactions in particular exhibited a stronger relationship to the physical distances and to the differences in preferred orientations than noise correlations. In contrast, the inferred negative interactions were less selective.

@@

Given a set of observations $\{x(t): t\in T$\} of population activity, where $x(t)$ is a $p\times 1$ vector of firing rates in time bin $t$, and an estimate $\bar x$ of the mean activity, the \emph{sample covariance matrix} is
\begin{equation}\label{eq:sample}
    C_{\sf sample} = \frac 1 n \sum\limits_{t\in T} (x(t)-\bar x)(x(t)-\bar x)^\T,
    \end{equation}
where $n$ is the number of time bins in $T$. 
The sample covariance matrix yields an unbiased estimate of the true covariance matrix ($\E{C_{\sf sample}}=\Sigma$) if the mean activity, $\bar x$, is estimated independently, in a separate data sample. 
In all cases when the unbiasedness of the sample covariance matrix matters in this paper, the means are estimated independently.

@@

The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}. Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of pairwise interactions \cite{Whittaker:1990}. For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}. More generally, partial correlations can serve as a measure of conditional independence under the assumption that dependencies in the system are close to linear effects \cite{Whittaker:1990,Baba:2004}. As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.

Pairwise partial correlations are closely related to the elements of the \emph{precision matrix}, \emph{i.e.}\;the inverse of the covariance matrix \cite{Dempster:1972,Whittaker:1990}. Zero elements in the precision matrix signify zero partial correlations between the corresponding pairs of variables. Given the covariance estimate $C$, the matrix of partial correlations $P$ is computed by normalizing the rows and columns of the \emph{precision matrix} $C^{-1}$ to produce negative unit entries on the diagonal:
\begin{equation}\label{eq:partial}
    P = -\left(\mbox{diag}(C^{-1})\right)^{-\frac 1 2} C^{-1} \left(\mbox{diag}(C^{-1})\right)^{-\frac 1 2}
\end{equation}

Increasing the number of recorded neurons results in a higher \emph{condition number} of the sample covariance matrix \cite{Ledoit:2004} making the partial correlation estimates more \emph{ill-conditioned}: small errors in the covariance estimates translate into greater errors in the estimates of the partial correlations. With massively multineuronal recordings, partial correlations cannot be estimated without \emph{regularization} \cite{Ledoit:2004,Schafer:2005}.

@@

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the sample covariance matrix $C_{\sf sample}$ toward $D$ as controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf sample} + \lambda D
\end{equation}
The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (Fig.~\ref{fig:1} Row 1, A).  If sample correlations are largely spurious, $C_{\sf diag}$ is expected to be more efficient than other estimators.

@@

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by the inverse of a sparse matrix $S$:
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \emph{i.e.}\;one in which a large fraction of off-diagonal elements are set to zero.  
The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$. 
The problem of finding the optimal set of non-zero elements of the precision matrix is known as \emph{covariance selection} \cite{Dempster:1972}. 
The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (`interactions') between some pairs of neurons (Fig.~\ref{fig:1} Row 1, C).

@@

\subsection*{Simulation}
We next demonstrated how the most efficient among different regularized estimators can reveal the structure of correlations.
We constructed four families of $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (Fig.~\ref{fig:1} Row 2, A--D and Methods).  We used these covariance matrices as the ground truth in multivariate Gaussian distributions with zero means and drew samples of various sizes. 
The sample correlation matrices from finite samples (\emph{e.g.}\ $n=500$ in Fig.~\ref{fig:1} Row 3) were contaminated with sampling noise and their underlying structures were difficult to discern.

The evaluation of any covariance matrix estimator, $C$, is performed with respect to a \emph{loss function} $\ell(C,\Sigma)$ to quantify its discrepancy from the truth, $\Sigma$.  The loss function is chosen to attain its minimum when $C=\Sigma$.
Here, in the role of the loss function we adopted the Kullback-Leibler divergence between multivariate normal distributions with equal means, scaled by $\frac 2 p$ to make its values comparable across different population sizes: 
\begin{equation}\label{eq:loss}
    \ell(C,\Sigma) = 
    \frac 2 p D_{KL}\left(\mathcal N_\Sigma\,\Vert\,\mathcal N_C\right) = 
    \frac 1 {p} \left[\Tr(C^{-1}\Sigma) + \ln\det C - \ln\det\Sigma - p\right] 
\end{equation}
Thus $\ell(C,\Sigma)$ is expressed in nats/neuron per time bin.

When the ground truth is not accessible, the loss cannot be computed directly but may be estimated from data through \emph{validation}.
In a validation procedure, a validation sample covariance matrix $C_{\sf sample}^\prime$ is computed from a testing data set that is independent from the data used for computing $C$.
Then the \emph{validation loss} $\loss{C,C^\prime_{\sf sample}}$ measures the discrepancy of $C$ from $C^\prime_{\sf sample}$. 
Here, in the role of validation loss, we adopted the negative multivariate normal log likelihood of $C$ given $C^\prime_{\sf sample}$, also scaled by $\frac 2 p$ and omitting the constant term: 
\begin{equation}\label{eq:vloss}
    \loss{C,C^\prime_{\sf sample}} = \frac 1 p \left[\Tr(C^{-1}C^\prime_{\sf sample})+\ln\det C \right]
\end{equation}

Since $\loss{\cdot,\cdot}$ is additive in its second argument and $C^\prime_{\sf sample}$ is an unbiased estimate of $\Sigma$, then, for given $C$ and $\Sigma$, the validation loss is an unbiased estimate of the true loss, up to a constant:
\begin{equation}
    \E{\loss{C,C_{\sf sample}^\prime}}=\loss{C,\E{C_{\sf sample}^\prime}}=\loss{C,\Sigma} = \ell(C,\Sigma) + \mbox{const}.
\end{equation}
Therefore, validation procedure allow comparing the relative values of the average loss attained by different covariance matrix estimators.

We drew 30 independent samples with sample sizes $n=250$, 500, 1000, 2000, and 4000 from each model and computed the loss $\ell(C,\Sigma)$ for each of the five estimators.  
The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  
All the regularized estimators produced better estimates (lower loss) than the sample covariance matrix.  
However, estimators whose structure matched the true model outperformed the other estimators (Fig.~\ref{fig:1} Row 5).
The validation loss computed by 10-fold cross-validation (see Methods) accurately reproduced the relative values of the true loss and the rankings of the covariance estimators even without access to the ground truth (Fig.~\ref{fig:1} Row 6).

@@

In Gaussian models, partial correlations perfectly characterize the conditional dependencies between variables and the graphical models of partial correlations exactly correspond to the conditional dependencies in the data. 
To demonstrate that estimator rankings were robust to deviations from Gaussian models, we repeated the same cross-validated evaluation using pairwise Ising models to generate the data.
Ising models have been used to infer functional connectivity from neuronal spike trains \cite{Hertz:2011}. 
Conveniently, the Ising model has equivalent mathematical form to the Gaussian distribution,
\begin{equation}\label{eq:ising}
    x\sim \frac 1 {Z(J,h)} \exp\left(\frac 1 2 x^\T J x + h^\T x \right)
\end{equation}
but the Ising model is defined on the multivariate binary domain rather than the continuous domain. 
Both models are maximum-entropy models constrained to match the mean firing rates and the covariance matrix \cite{Jaynes:1957}.
The partition function $Z(J,h)$ normalizes the distribution on the models' respective domains. 
In the Gaussian model, the matrix $-J^{-1}$ is the covariance matrix; and the mean values are $\mu=J^{-1}h$.  
For the Ising model, $J$ is the matrix of pairwise interactions and $h$ is the vector of the cells' individual activity drives, although they do not have a simple relationship to the means and the covariance matrix. 
Both distributions have the same structure of pairwise conditional dependencies: zeros in the matrix $J$ indicate conditional independence between the corresponding pair of neurons.

@@

The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:6} C). This suggests that correlations between neurons build up from multiple chains of smaller interactions. Furthermore, the average partial correlations were less variable ($p=0.002$ Brown-Forsythe test): the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29.

@@

\subsection*{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data can produce very large errors in the inferred network of dependencies (Fig.~\ref{fig:5} C and D). The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}. However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be practical or sufficient, and the problem must be addressed by using regularized estimators. Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.
